The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis

### 2024-04-23 carbon_paper/tex/main.tex
updated draft again

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,
% referee,
usenatbib]{mnras}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
% \usepackage{amssymb}	% Extra maths symbols

\usepackage{isotope}

\usepackage{hyperref}

\graphicspath{{figures/}}

@@

\newcommand{\citetjack}{Roberts et al.~(2023, in prep.)}
\newcommand{\citepjack}{(Roberts et al.~2023, in prep.)}
\newcommand{\citealtjack}{Robert et al.~2023, in prep.}
\newcommand{\cxi}{\texttt{\hyperlink{C11}{C11}}}
\newcommand{\kx}{\texttt{\hyperlink{K10}{K10}}}
\newcommand{\kxvi}{\texttt{\hyperlink{K16}{K16}}}
\newcommand{\vxiii}{\texttt{\hyperlink{V13}{V13}}}
\newcommand{\cfactor}{2.4}

% Acronyms
\newcommand{\agb}{AGB}
\newcommand{\apogee}{APOGEE}
\newcommand{\cc}{CCSNe}
\newcommand{\gce}{GCE}
\newcommand{\ia}{SNe Ia}
\newcommand{\imf}{IMF}
\newcommand{\sfh}{SFH} % check where first defined
\newcommand{\ssp}{SSP}

% internal abbreviations
\newcommand{\lims}{lower-intermediate mass stars}
\newcommand{\hms}{high-mass stars}
\newcommand{\caah}{[C/Mg]-[Mg/H]}
\newcommand{\caafe}{[C/Mg]-[Mg/Fe]}

@@

\newcommand{\Yct}{\ensuremath{y_{\rm C}}}
\newcommand{\Ycc}{\ensuremath{y_{\rm C}^{\rm cc}}}
\newcommand{\Yoc}{\ensuremath{y_{\rm Mg}^{\rm cc}}}
\newcommand{\Ycagb}{\ensuremath{y_{\rm C}^{\rm agb}}}
% \newcommand{\y}{\ensuremath{\rotatebox[origin=B,y=0.5ex]{180}{y}}}
\newcommand{\y}{Y}

\newcommand{\Mo}{%
    \ifmmode {\rm M}_{\sun}%
    \else {M$_{\sun}$}%
    \fi}
\newcommand{\Zo}{%
    \ifmmode Z_{\sun}%
    \else $Z_{\sun}$%
    \fi}

@@

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. A. Boyea et. al.]{
Daniel A. Boyea,$^{1}$\thanks{E-mail: boyea.2@osu.edu}
James W. Johnson,$^{1}$
Third Author$^{2,3}$
and Others$^{1,3}$
\\
% List of institutions
$^{1}$Department of Astronomy, the Ohio State University, 191 W. Woodruff, Columbus, OH 43210, USA
}

@@

% Enter the current year, for the copyright statements etc.
\pubyear{2023}

@@

% Abstract of the paper
\begin{abstract}
% context
C is an important element across astronomy; however, its origin remains poorly understood. 
% 
We aim to constrain the stellar yields of C through multi-zone Galactic chemical evolution models by comparing predictions with \apogee\ subgiants abundances.
% 
We find that \caafe\ is an emirical estimate of the delayed C sources, enabling us to estimate that \agb\ stars and \cc\ produce about 20\% and 80\% of C, respectively.  
The \caah\ trend instead represents the equilibrium abundances of C and Mg. 
We use the \caah\ trend to estimate the CCSNe C/Mg yield, determining that  $\Ycc/\Yoc = EQUATION$ when including \agb\ C. 
% misc
Our models are relatively independent of uniform scaling of yields and outflows, and alternate star formation histories. 
However, the stars which contribute to \agb\ C production and the \ia\ delay time distribution of Fe contribute uncertainties to our conclusions. 
% gas phase
While reliable gas-phase and low-metallicity measurments of C are challenging, we find that our model and a single-zone model with our recommended yields replicate the broad trends of \caah{} across different environments and metallicities.

@@

@@

Carbon is a distinctive and well-studied element in astronomy. 
Formed in the cores of stars during He fusion, C is the lightest directly synthesized element after He, one of the only light elements formed in low-mass stars, and one the most abundant metals \citep[e.g.][]{jennifer19, KL14}.\footnotemark{}
C structurally changes the environments it pollutes -- regulating stellar evolution, facilitating the formation of stars and planets, and forming the basis of earthly life.
% Because the effects of first dredge up are mass-dependent, C and N abundances are frequently used as age indicators for RGB stars \citep{MG15, martig16, hasselquist19, vincenzo+21}. 
As such, understanding the origin of C has wide ranging implications. 
While we know both lower-intermediate-mass and high-mass stars produce C, the relative importance of each process is still unknown.

\footnotetext{By metallicity, we mean the (mass) fraction of any element which is not H or He, denoted by $Z$. For the sun, we take $Z_\odot=0.014$. }

We know, from observations of MW stars and (extragalactic) gas, that C/O traces a banana shape in O/H (see Fig.~\ref{fig:gas_phase} and section~\ref{sec:gas}). 
At the very lowest metallicities, where $[{\rm O/H}]\lesssim-2$,%
\footnote{In this paper, we use the standard notation for chemical abundances. $[A/B] = \log_{10}\left(A/B\right) - \log_{10}\left(A_{\sun}/B_{\sun}\right)$, i.e. $[A/B]$ is the logarithm of the ratio between A and B, scaled such that $[A/B]=0$ for the sun. Solar abundances are as measured in \citet{asplund+09}.}
%
observations suggest that C/O declines with metallicity.
From $-2 \lesssim {\rm [O/H]}\lesssim -1$, C/O is roughtly constant with metallicities. 
And, at higher metallicities $-1 \lesssim {\rm [O/H]}$, C/O increases with increasing metallicity.
    As our primary observational constraint, we use a sample of subgiant stars from the \apogee{} \citep{apogee17} selected by the criteria in Roberts et al. (2023, in prep).
According to stellar evolution theory and observations, subgiants most accurately represent their birth composition \citep{gilroy89, korn+07, lind+08, souto+18, souto19}.
Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}. [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. 
Using the subgiant abundance trends, we will develop a model of the enrichment sources and evolution of C.

Galactic chemical evolution (\gce) is a powerful tool, capable of uncovering the origins of the elements. 
Each enrichment process has characteristic chemical signatures and timescales, enabling us to reconstruct chemical histories.
Many previous works have used GCE in attemt to understand C abundances, whether through using theoretical stellar models \citep{DTS78, prantzos+18, chiappini+03} 
or understanding  observation (\citealt{tinsley79, HEK00, BF06, rybizki+17, berg+19, KKL20};
See also review in \citealt{romano22}).
%
Every study agrees that C is produced by a combination of high-mass and \lims{}, different studies disagree on which process is dominant. 
For example CITES conclude \hms{} dominate wherease CITES conclude \lims{} contribute the majority of C.
C is also generally understood to have strong metallicity dependent CCSNe enrichment.

One of the primary uncertainties of GCE models are nucleosynthetic yields. Yield predictions -- the amounts of each chemical element stars produce --
are shaped by poorly understood processes, including mass loss, nuclear reaction rates, rotational mixing, convection, and explodability \citep{romano+10,KL14,ventura+13, LC18, emily+21}.
To better understand where C comes from and how it evolves, our aim is to combine \apogee\ observations and multi-zone models to develop observationally-consistant yields.
\cite{james+23} examined similar \gce{} models of N (which is closely related to C), finding that trends in N and O are explained by the metallicity dependence of N/O yields. \citet{james+23} determine that \agb\ N abundances roughly depend linearly on metallicity (i.e. $y_{\rm N}/y_{\rm O} \propto Z$). 
Here, we extend their models to C, deriving similar constraints on C/Mg yields. We assess which yield prescriptions reproduce Galactic abundance trends while investigating the impact of \gce{} model assumptions, such as the star formation history (\sfh{}) and outflow mass loading.

\section{Data Selection}

@@

Subgiants provide the ideal observational constraint to our model. 
When a star enters the Red Giant Branch (RGB), material from the CNO-processed core is mixed with the envelope in first dredge up, enhancing N and depleting C \citep{iben67, vincenzo+21,KL14}. RGB stars thus require model-dependent corrections to recover surface abundances \citep[e.g.][]{vincenzo+21}. On the other hand, gravitational settling can affect main sequence stellar abundances \citep[e.g.][]{souto19}. Subgiants have well-mixed envelopes, so gravitational settling is not as significant, and subgiants have not yet experienced first dredge up. We use a sample of \apogee\ DR17 stars \citep{apogee17} as selected in \citetjack.  
Chemical abundances are determined from the \apogee\ Stellar Parameter and Chemical Abundance Pipeline ({\sc aspcap}) \citep{aspcap}.

Fig.~\ref{fig:subgiant_selection} shows a plot of all \apogee\ stars and the \citetjack polygon selection criteria. 
 \citetjack~select a region of stars based on surface gravity $\log g$, and effective surface temperature, $T_\text{eff}$.
 \begin{equation}
    \begin{cases} \label{eq:subgiant_selection}
        \log \text{g} \geq 3.5 \\
        \log \text{g} \leq 0.004\,T_{\rm eff} - 15.7 \\
        \log \text{g} \leq 0.000706\,T_{\rm eff} + 0.36 \\
        \log \text{g} \leq -0.0015\,T_{\rm eff} + 12.05 \\
        \log \text{g} \geq 0.0012\,T_{\rm eff} - 2.8. \\
    \end{cases}
\end{equation}
Additionally, we exclude stars in \apogee{} marked by any of the following flags.
\begin{description}
\item \verb|APOGEE_MIRCLUSTER_STAR|
\item \verb|APOGEE_EMISSION_STAR|
\item \verb|APOGEE_EMBEDDEDCLUSTER_STAR|
\item \verb|young cluster (IN-SYNC)|
\item \verb|APOGEE2_W345|
\item \verb|EB planet|
\end{description}
Our final sample contains \about{12,000} subgiants.

When plotting the data and models, we bin by groups of 500 (or 150) stars in [Mg/H] ([Mg/Fe]). When plotting \caah, we select only the low-$\alpha$ sequence. For \caafe, we instead select stars with $-0.15 \leq {\rm [Mg/H]} \leq -0.05$.  
When plotting \caah, we isolate the low- and high-$\alpha$ sequences with the cut
\begin{equation}\label{eq:high_alpha}
\begin{cases}
\text{[Mg/Fe]} >0.12-0.13\,\text{[Fe/H]}, & \text{[Fe/H]}<0\\
\text{[Mg/Fe]} >0.12, & \text{[Fe/H]}>0. \\
\end{cases}
\end{equation}
The low-$\alpha$ sequence is better reproduced by this model, so we use this cut of the subgiants to compare the models against except for comparing \caafe.

\begin{figure}
    \centering
    \includegraphics{logg_jack.pdf}
    \caption[]{
        A Kiel diagram of \apogee{} stars. Following \citetjack, we select subgiants in the black pentagon (see Equation \ref{eq:subgiant_selection}). These stars have not yet experienced first dredge-up, so their photospheric C and N abundances should reflect their birth mixture.
    }
    \label{fig:subgiant_selection}
\end{figure}

\section{Nucleosynthesis}
Table \ref{tab:fiducial_mod} contains our fiducial yields. 
For O, N, Mg, and Fe, we adapt the yield choices of \citet{james+21, james+23}.
Following \citet{james+21, james+23}, we take the \ia{} delay time distribution to be a $t^{-1.1}$ power-law with a minimum delay time of 140\,Myr, as suggested by the observations of \citet{maoz+12}.

\begin{table}
	\centering
    \caption[]{Yields for the fiducial model (in units of \ssp~birth mass). See section \ref{sec:agb} for the definition of \cxi.}
	\label{tab:fiducial_mod}

	\begin{tabular}{l l l l}
		\hline
        Element & $y^{\rm cc}$ & $\y^{\rm agb}$ & $y^{\rm ia}$ \\
		\hline
        C & Eq.~\ref{eq:zeta} & $\cfactor\times$\cxi &  0 \\
        O & 0.015 & 0 & 0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

    In this work, we explore four different sets of \agb\ star yield tables from literature,
    providing necessary well-sampled grids in mass and metallicity. We refer to the yields from the following studies as the following,
\begin{description}
    \item \hypertarget{C11}{\texttt{C11}}: \citet{cristallo+11, cristallo+15}
    \item \hypertarget{K10}{\texttt{K10}}: \citet{karakas10}
\item \hypertarget{V13}{\texttt{V13}}: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \hypertarget{K16}{\texttt{K16}}: \citet{KL16, karakas+18}
\end{description}
Table~\ref{tab:agb} contains the masses and metallicities used in each model.

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yields for these four models.
We define the stellar \agb\ yield to be the fraction of the stars initial mass $M$ which is newly synthesized into C. If $\Delta Z_{\rm C}$ represents the change in the mass fraction of C abundance, then
\begin{equation}
    \y_{\rm C}^{\rm agb} =  \Delta Z_{\rm C} \frac{M_{\rm ejected}}{M}.
\end{equation}
Note that yields may be negative if the material returned to the interstellar medium has a lower abundance $Z_{\rm C}$ than the material the star was formed from.
Most models agree on the qualitative shape of the net fractional \agb\ C yield.
Stellar yields peak between masses of about 2--4 \Mo and decline as stars become more or less massive. As metallicity increases, the total \agb\ C yield decreases. The mass of peak C yields also increases slightly with metallicity.

The left panel of Fig.~\ref{fig:agb-ssp} shows \imf-averaged C yields for each \agb\ model as a function of metallicity.
An IMF-averaged yield adds together the yields of stars of each mass, weighted by the fraction of stars of each mass (the \imf). 
\begin{equation} \label{eq:imf-agb-yield}
    \Ycagb(Z,t) = 
    \frac{
    \int_{M_{\rm to}(t)}^{8\,\Mo} 
    \y_{\rm X}(M, Z)
    \frac{dN}{dM}\ M\ dM
}
{
    \int_{0.08\Mo}^{100\Mo} \frac{dN}{dM}\ M\ dM
}
\end{equation}
where ${dN}/{dM}$ is the \imf{} and $M_{\rm to}(t)$ is the mass of stars with lifetime $t$.%
\footnotetext{In our model, the mass-lifetime relation is
$\log \tau_M = 1.02 - 3.57\log M + 0.90 \left(\log M\right)^2$,
where $\tau_M$ is in Gyr, from \citealt{larson74}.
We use $t_{\rm end}=10\,$Gyr for total yields when $t$ is not used.}
The \agb\ models mostly differ in their yield normalization and metallicity dependence. 
All yield models span a range of \about{2} for a given metallicity.
For example, all four models predict $y_\text{C}^\text{agb}(\Zo)$ to be between 0.004 and 0.008 at solar metallicity. The models also differ slightly in the strength of metallicity dependence, with about a factor of 3 discrepency (see Table~\ref{tab:agb}).

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,
%referee, % this line changes to double-spaced 1 column
usenatbib]{mnras}

@@

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
% \usepackage{amssymb}	% Extra maths symbols
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\usepackage[dvipsnames]{xcolor}
% \usepackage{enumitem}

\graphicspath{{figures/}} 
% \linespread{1.8}

@@

\newcommand{\citetjack}{\citet{jack}}
\newcommand{\citepjack}{(Roberts et al.~2024, in prep.)}
\newcommand{\citealtjack}{Robert et al.~2024, in prep.}
\newcommand{\cxi}{\texttt{\hyperlink{C11}{C11}}}
\newcommand{\pxvi}{\texttt{\hyperlink{P16}{P16}}}
\newcommand{\kxvi}{\texttt{\hyperlink{K16}{K16}}}
\newcommand{\vxiii}{\texttt{\hyperlink{V13}{V13}}}
\newcommand{\cfactor}{2.4}
\newcommand{\nsubgiants}{16,365}

% Acronyms
\newcommand{\agb}{AGB}
\newcommand{\apogee}{APOGEE}
\newcommand{\aspcap}{\textsc{aspcap}}
\newcommand{\cc}{CCSN}
\newcommand{\gce}{GCE}
\newcommand{\ia}{SN Ia}

% internal abbreviations
\newcommand{\caah}{[C/Mg]-[Mg/H]}
\newcommand{\caafe}{[C/Mg]-[Mg/Fe]}

@@

\newcommand{\Yct}{\ensuremath{y_{\rm C}}}
\newcommand{\Ycc}{\ensuremath{y_{\rm C}^{\rm CC}}}
\newcommand{\Yoc}{\ensuremath{y_{\rm Mg}^{\rm CC}}}
\newcommand{\Ycagb}{\ensuremath{y_{\rm C}^{\rm AGB}}}
% \newcommand{\y}{\ensuremath{\rotatebox[origin=B,y=0.5ex]{180}{y}}}
\newcommand{\y}{Y}

\newcommand{\Mo}{%
    \ifmmode {\rm M}_{\sun}%
    \else {M$_{\sun}$}%
    \fi}

\newcommand{\Zo}{%
    \ifmmode Z_{\sun}%
    \else $Z_{\sun}$%
    \fi}

@@

%%% citepos command
\makeatletter
\DeclareRobustCommand\citepos
  {\begingroup
   \let\NAT@nmfmt\NAT@posfmt% ...except with a different name format
   \NAT@swafalse\let\NAT@ctype\z@\NAT@partrue
   \@ifstar{\NAT@fulltrue\NAT@citetp}{\NAT@fullfalse\NAT@citetp}}

\let\NAT@orig@nmfmt\NAT@nmfmt
\def\NAT@posfmt#1{\NAT@orig@nmfmt{#1's}}

\makeatother

%%%%%%%%%%% JWJ' editing macros %%%%%%%%%%%
\newcommand{\strike}[1]{{\color{ForestGreen} \sout{#1}}}
\newcommand{\add}[1]{{\color{ForestGreen} #1}}
\newcommand{\note}[1]{{\color{ForestGreen} \textit{ \small (JWJ: #1)}}}
% \LetLtxMacro\origcitep\citep
% \LetLtxMacro\origcitet\citet
% \renewcommand{\citep}[2][]{%
%     \mbox{\origcitep[#1][]{#2}}
% }
% \renewcommand{\citet}[2][]{%
%     \mbox{\origcitet[#1][]{#2}}
% }

\newcommand{\dbstrike}[1]{{\color{Thistle} \sout{#1}}}
\newcommand{\dbadd}[1]{{\color{Thistle} #1}}
\newcommand{\dbnote}[1]{{\color{Thistle} \textit{\small (DB: #1)}}}

% \LetLtxMacro\origcite\cite

% This is my command. I want to strike out, 
% \newcommand{\strikeit}[1]{%
%   \ifcorrectingmode
%   \mbox{\sout{#1}}%
% %  \makebox[\textwidth][s]{\sout{#1}}%
%   \fi
% }

% \renewcommand{\cite}[2][]{%
%   \ifcorrectingmode
%   \mbox{\origcite[#1]{#2}}%
%   \else
%   \origcite[#1]{#2}%
%   \fi
% }

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. A. Boyea et. al.]{%
Daniel A. Boyea,$^{1, 2, 3}$\thanks{%
Contact e-mail:~\href{mailto:boyea.2@osu.edu}{boyea.2@osu.edu}}
James W. Johnson,$^{1, 2, 4}$
Third Author$^{2,3}$
and Others$^{1,3}$
\\
% List of institutions
$^{1}$Department of Astronomy, The Ohio State University, 140 W. 18th Ave., Columbus, OH, 43210, USA
\\
$^{2}$Center for Cosmology \& Astroparticle Physics (CCAPP), The Ohio State University, 191 W. Woodruff Ave., Columbus, OH, 43210, USA
\\
$^{3}$Department of Physics \& Astronomy, University of Victoria, Victoria, BC V8P 5C2, Canada
\\
$^{4}$The Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena, CA, 91101, USA
}

@@

% Enter the current year, for the copyright statements etc.
\pubyear{2024}

@@

% Abstract of the paper
\begin{abstract}
% context
The origin of C, despite its importance, remains poorly understood. 
% 
We aim to constrain the stellar yields of C through multi-zone Galactic chemical evolution models by comparing predictions with \apogee\ subgiants abundances.
% 
We find that \caafe\ trends, when restricted in metallicity, are an empirical estimate of the delayed C sources, enabling us to estimate that \agb\ stars and \cc\ produce about 20\% and 80\% of C, respectively.  
The \caah\ trend instead represents the equilibrium abundances of C and Mg. 
We use the \caah\ trend to estimate the CCSNe C/Mg yield, determining that  $\Ycc/\Yoc = EQUATION$ when including \agb\ C. 
% misc
Our models are relatively independent of uniform scaling of yields and outflows, and alternate star formation histories. 
However, the stars which contribute to \agb\ C production and the \ia\ delay time distribution of Fe contribute uncertainties to our conclusions. 
% gas phase
While reliable gas-phase and low-metallicity measurements of C are challenging, we find that our model and a single-zone model with our recommended yields replicate the broad trends of \caah{} across different environments and metallicities.

@@

\begin{keywords}
galaxies: abundances -- galaxies: evolution -- galaxies: star formation -- galaxies: stellar content -- methods: numerical
\end{keywords}

@@

Despite being one of the most abundant metals in the universe~\citep[e.g.][]{asplund+09}, C nucleosynthesis is poorly understood.
Formed in the cores of stars during He fusion, C is the first \dbadd{primary? element} after He. Additionally, C is one of the only light elements formed in low-mass stars \citep[e.g.,][]{KL14}. \dbnote{I don't like how many e.g.'s there are, and is this the right citation here?}
\dbadd{
Because the effects of first dredge up are mass-dependent, C and N abundances are frequently used as age indicators for RGB stars \citep{MG15, martig16, hasselquist19, vincenzo+21}. 
}
\dbnote{C abundnces for stellar structure (opacities) and star formation (line cooling)}
Broadly, it is well understood that both lower-intermediate-mass and high-mass stars produce C. However, the relative yields from these two classes of stars and their dependence on  metallicitys unknown \add{(see discussion below)}.

% \footnotetext{\strike{By metallicity, we mean the (mass) fraction of any element which is not H or He, denoted by $Z$. For the sun, we take $Z_\odot=0.016$. CHECK THIS~} }

Like many other metals, C nucleosynthesis \dbstrike{is poorly understood because its predicted yields} \dbadd{predictions from stellar evolution are highly uncertain \citep[see reviews][]{romano+10, KL14}}.
\strike{Yield predictions} are shaped by poorly understood processes, such as mass loss rates \citep{ventura+13}?, nuclear reaction networks, rotational mixing \citep{frischknecht+16, LC18}, and convection ~\citep{chieffi2001}. \dbnote{check citations here...}. 
Binarity (and the metallicity dependence thereof), e.g. MKB2019, farme,
Small changes in the initial mass ($0.01\Mo$) of massive stars can dramatically change the subsequent evolution \citep{bruenn+2023, emily+21}. As such, accurate yield predictions would require a much more densly sampled set of models in mass, metallicity, and rotation rates then are presently available. Furthermore, binary evolution is rarely included in massive star models with possibly important implications for yields \citep{farmer+21}, despite the large (metallicity-dependent) fraction of massive stars which live in binary systems \citep{MKB2019}.

\dbnote{added paragraph here}
In light of these difficulties with accurate theoretical modeling, recent work has instead sought to characterize chemical yields empirically.
\citet{HLA20} measured the abundance ratios in the ejecta of Cassiopeia A using spatial resolved X-ray spectra.
\citet{weinberg+19, weinberg+22} analyzed the abundance ratio trends of high- and low-alpha sequence stars in APOGEE~\citep{apogee17} to infer relative yields of prompt and delayed enrichment sources and the metallicity dependence thereof.
\citet{emily+19, emily+22, emily+23} applied the same methodology to GALAH~\citep{DeSilva2015, Martell2017}.
\citet{rodriguez+21, rodriguez+23} measured the mean Fe yield of core-collapse supernovae (CCSNe) using the radioactive tails of their lightcurves.

\dbnote{combine this with previous, shorten J23 description and call out other ongoing work e.g. Miquela's He, James's chemical equilibrium, ...}
Most notably for our purposes,~\citet{james+23} used Galactic chemical evolution (GCE) models to show that the correlation between gas-phase N and O abundances (see their Fig. 1) can be explained by the metallicity dependence of relative N and O yields. 
Variations in assumptions regarding GCE parameters, such as the detailed star formation history (SFH), did not impact their conclusions and are instead a likely source of scatter in the observed relation.
Here, we apply similar methodology to constrain C yields, which are closely related to N.

GCE connects a galaxy's evolutionary history to the detailed elemental abundances in its gas and stars.
Each enrichment process has characteristic chemical signatures and timescales, enabling us to reconstruct \strike{chemical} \add{enrichment} histories.
\add{As an element that is well-measured in stellar spectra, many} \dbadd{with large samples of stellar C abundancess}, \strike{Many} previous GCE works \add{have} focused on C abundances (\citealt{DTS78, BF06, prantzos+18, berg+19}; see also the recent reviews by \citealt{romano22} and \citealt{RM21}).
To date there is no consensus on the detailed origin of C; some authors have argued that high-mass stars should dominate~\citep[e.g.][]{prantzos+94, HEK00, romano+20, franchini+20, gustafsson22}, and others have argued that AGB stars should contribute at least half of the C in the universe~\citep[e.g.][]{tinsley79, chiappini+03, mattsson10, KKU11, rybizki+17, KKL20}.

    As our primary observational constraint, we use a sample of subgiant stars from the \apogee{} \citep{apogee17} selected by the criteria in \citetjack.
According to stellar evolution theory and observations, subgiants most accurately represent their birth composition \citep{gilroy89, korn+07, lind+08, souto+18, souto19}.
Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}.\footnotemark{} In this sample, [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. 
From observations of nearby stars and (extragalactic) gas, we know that [C/O] against [O/H] traces a banana shape (see Fig.~\ref{fig:gas_phase} and section~\ref{sec:gas}).
At the very lowest metallicities\strike{, where} \add{(}$[{\rm O/H}]\lesssim-2$\add{)},%
%
observations \strike{suggest} \add{indicate} that [C/Mg] declines with metallicity.
From $-2 \lesssim {\rm [O/H]}\lesssim -1$, [C/Mg] is roughly constant with metallicity. 
At higher metallicities ([O/H]~$\gtrsim -1$), [C/Mg] increases with metallicity.

\strike{Using the subgiant abundance trends, we will develop a model of the enrichment sources and evolution of C.}
\add{We will show in section~\ref{sec:results} below that these correlations in the bulk population abundances alone are quite constraining for the origin of C.}

\footnotetext{In this paper, we use the standard notation for chemical abundances. $[A/B] = \log_{10}\left(A/B\right) - \log_{10}\left(A_{\sun}/B_{\sun}\right)$, i.e. $[A/B]$ is the logarithm of the mass ratio between A and B, scaled such that $[A/B]=0$ for the sun (see Table~\ref{tab:fiducial_mod} for solar scale.)}

In Section~\ref{sec:data_selection} we describe the selection for our subgiant observational sample.
Section~\ref{sec:nucleosynthesis} discusses theoretical estimates of \agb\ and \cc\ C yields. We parameterize our yield choices and derive a estimate of C/Mg yields observationally.
Section~\ref{sec:vice} reviews the multizone model and modifications thereof for this work.
Section~\ref{sec:results} presents our multizone model predictions and GCE parameters which are observationally permissable. 
Finally Section~\ref{sec:gas} compares our model to literature collected observations of halo stars and extragalactic HII regions.

\section{Data Selection}\label{sec:data_selection}

@@

% Subgiants are a useful empirical benchmark because their photospheric abundances should reflect their birth mixture. 
As an emperical benchmark, we use subgiant stars, whose photospheres accurately represent birth abundances. 
During main sequence evolution, metals can fall to below the convective envelope (i.e., ``gravitational settling").
When stars leave the main sequence and become subgiants, these metals are mixed with the convective envelope \citep[]{gratton+00, souto19}. 
 \textit{First dredge up} during the red giant phase brings CNO processed core material to the surface \citet{iben67, KL14}.

We use the sample of \apogee\ DR17 subgiants from \citetjack, selecting stars within the $\log g$-$T_\text{eff}$ polygon given by
 \begin{equation}
    \begin{cases} \label{eq:subgiant_selection}
        \log \text{g} \geq 3.5 \\
        \log \text{g} \leq 0.004\,T_{\rm eff} - 15.7 \\
        \log \text{g} \leq 0.000706\,T_{\rm eff} + 0.36 \\
        \log \text{g} \leq -0.0015\,T_{\rm eff} + 12.05 \\
        \log \text{g} \geq 0.0012\,T_{\rm eff} - 2.8. \\
    \end{cases}
\end{equation}
Also following \citet{jack}, we exclude stars with the flags:
        \verb|APOGEE_MIRCLUSTER_STAR|,
        \verb|APOGEE_EMISSION_STAR|,
        \verb|APOGEE_EMBEDDEDCLUSTER_STAR|,
        \verb|APOGEE2_YOUNG_CLUSTER|,
            \verb|APOGEE2_W345|,
        \verb|APOGEE2_EB|, and
        \verb|ASPCAP_BAD|.
We additionally remove stars with no reported C, Mg, or Fe abundances. These cuts yield a sample containing \nsubgiants\ subgiants.

\section{Nucleosynthesis}\label{sec:nucleosynthesis}

Table \ref{tab:fiducial_mod} contains our fiducial yields and abundance scale. 
For O, Mg, and Fe, we adopt the yields in \citet{david_fe}. We use a scaled variation of the N yield from \citet{james+23}. 
Following \citet{james+21, james+23}, we take the \ia{} delay time distribution to be a $t^{-1.1}$ power-law with a minimum delay time of 150\,Myr, as suggested by the observations of \citet{maoz+12}. We use solar abundances from \citet{magg+22} with a +0.04 gravitational settiling correction as in \citet{david_fe}. \dbadd{Magg22 finds Z/X = 0.0225 for the surface? what do we use for the solar metallicity?}. Assuming He abundance from $Y=0.2725$ from \citet{AAG2021}, resulting in a $Z_\odot=0.01606$

\begin{table}
	\centering
    \caption[]{Solar abundance scale and fiducial yields (in units of SSP~birth mass). See section \ref{sec:agb} for the definition of \cxi. The solar abundance scale is \citet{magg+22} + 0.04. }
	\label{tab:fiducial_mod}

	\begin{tabular}{l l l l l}
		\hline
        Element & $Z_{X,\,\sun}$ & $y_X^{\rm cc}$ & $\y_X^{\rm agb}$ & $y_X^{\rm ia}$  \\
		\hline
        C & 0.00339 & Eq.~\ref{eq:y_cc} & $\cfactor\times$\cxi &  0 \\
        O & 0.00733 & 7.12e-3 & 0 & 0 \\
        Mg & 0.000671 & 6.69e-4 & 0 & 0 \\
        Fe & 0.00137 & 4.72e-4 & 0 & 6.69e-4 \\
        N &0.00104 & 4.001e-4 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
        % C & 0.00339 & Eq.~\ref{eq:zeta} & $\cfactor\times$\cxi &  0 \\
        % O & 0.00733 & 0.015 & 0 & 0 \\
        % Mg & 0.000671 & 0.00185 & 0 & 0 \\
        % Fe & 0.00137 & 0.0012 & 0 & 0.00214 \\
        % N &0.00104 & 0.00072 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

\dbnote{I like the definition of Y to go here since that is model independent, but IMF-integrated could still be specific to AGB/CC.}

@@

An Asymptotic Giant Branch (AGB) star is a low mass star during its last nuclear burning phase (He shell burning; see e.g. \citealt{PR2023}) which enriches the ISM through mass loss during thermal pulses.
    In this work, we explore four different stellar AGB yield tables from the literature which provide
    well-sampled grids in mass and metallicity. We refer to the tables as 
\begin{description}
    \item \hypertarget{C11}{\texttt{C11}}: \citet{cristallo+11, cristallo+15}
    \item \hypertarget{V13}{\texttt{V13}}: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \hypertarget{K16}{\texttt{K16}}: \citet{KL16, karakas+18}
    \item \hypertarget{P16}{\texttt{P16}}: \citet{pignatari+16, ritter+18, battino+19, battino+21}
\end{description}
Table~\ref{tab:agb} notes the masses and metallicities for each grid of yields.
We use the same tables as \citet{james+23}, except that we have swapped \citet{karakas10} for \citet{ritter+18}.

We define a yield to be the fraction of a star's initial mass $M_{\rm birth}$ which is converted into freshly synthesized C. If $\Delta Z_{\rm C}$ represents the change in the mass fraction of C abundance, then
\begin{equation}
    \y_{\rm C}^{\rm AGB} =  \Delta Z_{\rm C} \frac{M_{\rm ejected}}{M_{\rm birth}}.
\end{equation}
The yields may be negative if the material returned to the ISM has a lower $Z_{\rm C}$ than the material the star was formed from.

Fig.~\ref{fig:y_agb} compares the stellar \agb\ C yields for these four models.
Most models agree on general yield trends in mass and metallicity.
C production peak between masses of $\sim$ 2--4 \Mo\ and declines as stars become more or less massive. As metallicity increases, the net yield decreases and the mass of peak C production increases slightly.
